---
title: "Tune machine learning models with MELD + additional variables"
output: html_document
date: "2024-08-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(compareC)
library(kableExtra)
library(gtsummary)

library(survival)
library(glmnet)
library(randomForestSRC)
library(gbm)
library(mboost)
library(xgboost)
library(CoxBoost)
library(survivalmodels)

library(glmnetUtils)
library(mlr3)
library(mlr3proba)
library(paradox)
library(mlr3tuning)
library(mlr3extralearners)
library(mlr3pipelines)

# Set background to be white for all ggplots
theme_set(theme_classic())

# Parallelization for CoxBoost
library(snowfall)
sfInit(parallel = TRUE, cpus = parallel::detectCores() - 2)
```

## Data processing

Read in the training data and two external validation datasets. 

```{r}
dir = "/Users/jliang/Library/CloudStorage/Box-Box/MELD\ QSU/data/"
train_df = read.csv(paste0(dir, "MELD3_Dev_Cohort.csv"))
test1_df = read.csv(paste0(dir, "MELD3_Val_Cohort.csv"))
test2_df = read.csv("/Users/jliang/Library/CloudStorage/Box-Box/STAR_files_as_of_4_5_2024/MELD\ 3.0/MELD3_SecVal_Cohort_2019_23.csv")
```

Subset the data to only include the covariates of interest (MELD-transformed and raw variables) and time-to-event for 90-day death. Split the second test set into two: 1/1/2019 to 1/1/2021 (exclusive) and 1/1/2021 (inclusive) to end of available data. 

```{r}
# Continuous and categorical variables
my_cont_vars = c("log_MELD_BILIRUBIN", "INIT_BILIRUBIN", 
                 "MELD_SERUM_SODIUM", "INIT_SERUM_SODIUM", 
                 "log_MELD_INR", "INIT_INR", 
                 "log_MELD_SERUM_CREAT", "INIT_SERUM_CREAT", 
                 "MELD_ALBUMIN", "INIT_ALBUMIN", 
                 "log_AGE", "INIT_AGE")
my_cat_vars = c("GENDER", "INIT_DIALYSIS_PRIOR_WEEK", 
                "Ethnicity_Cat", "ASCITES_BL", "INIT_ENCEPH")
my_vars = c(my_cont_vars, my_cat_vars)

my_train_df = train_df %>% 
  mutate(
    # Convert categorical variables to factors
    mutate(across(all_of(my_cat_vars), ~factor(.))), 
    # Log transform age
    log_AGE = log(INIT_AGE), 
    # Set lower/upper bounds for MELD variables and log transform
    MELD_BILIRUBIN = ifelse(INIT_BILIRUBIN < 1, 1, INIT_BILIRUBIN), 
    log_MELD_BILIRUBIN = log(MELD_BILIRUBIN), 
    MELD_SERUM_SODIUM = 137 - ifelse(INIT_SERUM_SODIUM > 137, 137, 
                                     ifelse(INIT_SERUM_SODIUM < 125, 125, 
                                            INIT_SERUM_SODIUM)), 
    MELD_INR = ifelse(INIT_INR < 1, 1, INIT_INR), 
    log_MELD_INR = log(MELD_INR), 
    MELD_SERUM_CREAT = ifelse(INIT_SERUM_CREAT > 3 | INIT_DIALYSIS_PRIOR_WEEK == "Y", 3, 
                              ifelse(INIT_SERUM_CREAT < 1, 1, INIT_SERUM_CREAT)), 
    log_MELD_SERUM_CREAT = log(MELD_SERUM_CREAT), 
    MELD_ALBUMIN = 3.5 - ifelse(INIT_ALBUMIN > 3.5, 3.5, 
                                ifelse(INIT_ALBUMIN < 1.5, 1.5, INIT_ALBUMIN))) %>% 
  select(all_of(c(my_vars, "SurvTime_by90d", "Death_by90d"))) %>% 
  na.omit()

my_test1_df = test1_df %>% 
  mutate(
    # Convert categorical variables to factors
    mutate(across(all_of(my_cat_vars), ~factor(.))), 
    # Log transform age
    log_AGE = log(INIT_AGE), 
    # Set lower/upper bounds for MELD variables and log transform
    MELD_BILIRUBIN = ifelse(INIT_BILIRUBIN < 1, 1, INIT_BILIRUBIN), 
    log_MELD_BILIRUBIN = log(MELD_BILIRUBIN), 
    MELD_SERUM_SODIUM = 137 - ifelse(INIT_SERUM_SODIUM > 137, 137, 
                                     ifelse(INIT_SERUM_SODIUM < 125, 125, 
                                            INIT_SERUM_SODIUM)), 
    MELD_INR = ifelse(INIT_INR < 1, 1, INIT_INR), 
    log_MELD_INR = log(MELD_INR), 
    MELD_SERUM_CREAT = ifelse(INIT_SERUM_CREAT > 3 | INIT_DIALYSIS_PRIOR_WEEK == "Y", 3, 
                              ifelse(INIT_SERUM_CREAT < 1, 1, INIT_SERUM_CREAT)), 
    log_MELD_SERUM_CREAT = log(MELD_SERUM_CREAT), 
    MELD_ALBUMIN = 3.5 - ifelse(INIT_ALBUMIN > 3.5, 3.5, 
                                ifelse(INIT_ALBUMIN < 1.5, 1.5, INIT_ALBUMIN))) %>% 
  select(all_of(c(my_vars, "SurvTime_by90d", "Death_by90d"))) %>% 
  na.omit()

my_test2_df = test2_df %>% 
  # 1/1/2019 to 1/1/2021 (exclusive)
  filter(mdy(INIT_DATE) < as.Date("2021-01-01")) %>% 
  mutate(
    # Convert categorical variables to factors
    mutate(across(all_of(my_cat_vars), ~factor(.))), 
    # Log transform age
    log_AGE = log(INIT_AGE), 
    # Set lower/upper bounds for MELD variables and log transform
    MELD_BILIRUBIN = ifelse(INIT_BILIRUBIN < 1, 1, INIT_BILIRUBIN), 
    log_MELD_BILIRUBIN = log(MELD_BILIRUBIN), 
    MELD_SERUM_SODIUM = 137 - ifelse(INIT_SERUM_SODIUM > 137, 137, 
                                     ifelse(INIT_SERUM_SODIUM < 125, 125, 
                                            INIT_SERUM_SODIUM)), 
    MELD_INR = ifelse(INIT_INR < 1, 1, INIT_INR), 
    log_MELD_INR = log(MELD_INR), 
    MELD_SERUM_CREAT = ifelse(INIT_SERUM_CREAT > 3 | INIT_DIALYSIS_PRIOR_WEEK == "Y", 3, 
                              ifelse(INIT_SERUM_CREAT < 1, 1, INIT_SERUM_CREAT)), 
    log_MELD_SERUM_CREAT = log(MELD_SERUM_CREAT), 
    MELD_ALBUMIN = 3.5 - ifelse(INIT_ALBUMIN > 3.5, 3.5, 
                                ifelse(INIT_ALBUMIN < 1.5, 1.5, INIT_ALBUMIN))) %>% 
  select(all_of(c(my_vars, "SurvTime_by90d", "Death_by90d"))) %>% 
  na.omit()

my_test3_df = test2_df %>% 
  # 1/1/2021 (inclusive) to end of available data
  filter(mdy(INIT_DATE) >= as.Date("2021-01-01")) %>% 
  mutate(
    # Convert categorical variables to factors
    mutate(across(all_of(my_cat_vars), ~factor(.))), 
    # Log transform age
    log_AGE = log(INIT_AGE), 
    # Set lower/upper bounds for MELD variables and log transform
    MELD_BILIRUBIN = ifelse(INIT_BILIRUBIN < 1, 1, INIT_BILIRUBIN), 
    log_MELD_BILIRUBIN = log(MELD_BILIRUBIN), 
    MELD_SERUM_SODIUM = 137 - ifelse(INIT_SERUM_SODIUM > 137, 137, 
                                     ifelse(INIT_SERUM_SODIUM < 125, 125, 
                                            INIT_SERUM_SODIUM)), 
    MELD_INR = ifelse(INIT_INR < 1, 1, INIT_INR), 
    log_MELD_INR = log(MELD_INR), 
    MELD_SERUM_CREAT = ifelse(INIT_SERUM_CREAT > 3 | INIT_DIALYSIS_PRIOR_WEEK == "Y", 3, 
                              ifelse(INIT_SERUM_CREAT < 1, 1, INIT_SERUM_CREAT)), 
    log_MELD_SERUM_CREAT = log(MELD_SERUM_CREAT), 
    MELD_ALBUMIN = 3.5 - ifelse(INIT_ALBUMIN > 3.5, 3.5, 
                                ifelse(INIT_ALBUMIN < 1.5, 1.5, INIT_ALBUMIN))) %>% 
  select(all_of(c(my_vars, "SurvTime_by90d", "Death_by90d"))) %>% 
  na.omit()


# Convert multi-level factors to indicators
my_train_df = data.frame(my_train_df, 
                         model.matrix(~Ethnicity_Cat + ASCITES_BL + INIT_ENCEPH, 
                                      my_train_df)[,-1])
my_test1_df = data.frame(my_test1_df, 
                         model.matrix(~Ethnicity_Cat + ASCITES_BL + INIT_ENCEPH, 
                                      my_test1_df)[,-1])
my_test2_df = data.frame(my_test2_df, 
                         model.matrix(~Ethnicity_Cat + ASCITES_BL + INIT_ENCEPH, 
                                      my_test2_df)[,-1])
my_test3_df = data.frame(my_test3_df, 
                         model.matrix(~Ethnicity_Cat + ASCITES_BL + INIT_ENCEPH, 
                                      my_test3_df)[,-1])

my_factor_vars = list(
  "Ethnicity_Cat" = c("Ethnicity_Cat2_Black", "Ethnicity_Cat4_Hispanic", 
                      "Ethnicity_Cat5_Asian", "Ethnicity_Cat6_Other"), 
  "ASCITES_BL" = c("ASCITES_BLModerate", "ASCITES_BLSlight"), 
  "INIT_ENCEPH" = c("INIT_ENCEPH2", "INIT_ENCEPH3")
)


# Set up tasks for each dataset
task_train = TaskSurv$new(
  "my_train_df",
  my_train_df %>% 
    mutate(GENDER = ifelse(GENDER == "F", 1, 0), 
           INIT_DIALYSIS_PRIOR_WEEK = ifelse(INIT_DIALYSIS_PRIOR_WEEK == "Y", 1, 0)) %>% 
    dplyr::select(-names(my_factor_vars)), 
  time = "SurvTime_by90d", event = "Death_by90d")
task_test1 = TaskSurv$new(
    "my_test1_df", 
    my_test1_df %>% 
      mutate(GENDER = ifelse(GENDER == "F", 1, 0), 
             INIT_DIALYSIS_PRIOR_WEEK = ifelse(INIT_DIALYSIS_PRIOR_WEEK == "Y", 1, 0)) %>% 
    dplyr::select(-names(my_factor_vars)), 
    time = "SurvTime_by90d", event = "Death_by90d")
task_test2 = TaskSurv$new(
    "my_test2_df", 
    my_test2_df %>% 
      mutate(GENDER = ifelse(GENDER == "F", 1, 0), 
             INIT_DIALYSIS_PRIOR_WEEK = ifelse(INIT_DIALYSIS_PRIOR_WEEK == "Y", 1, 0)) %>% 
    dplyr::select(-names(my_factor_vars)), 
    time = "SurvTime_by90d", event = "Death_by90d")
task_test3 = TaskSurv$new(
    "my_test3_df", 
    my_test3_df %>% 
      mutate(GENDER = ifelse(GENDER == "F", 1, 0), 
             INIT_DIALYSIS_PRIOR_WEEK = ifelse(INIT_DIALYSIS_PRIOR_WEEK == "Y", 1, 0)) %>% 
    dplyr::select(-names(my_factor_vars)), 
    time = "SurvTime_by90d", event = "Death_by90d")
```

## Modeling

### MELD

MELD3 score. 

```{r}
# Predict on data sets
preds_MELD_train = my_train_df %>% 
  mutate(
    MELD3 = round(1.33 * ifelse(GENDER == "F", 1, 0) + 
                    4.56 * log_MELD_BILIRUBIN + 0.82 * MELD_SERUM_SODIUM - 
                    0.24 * MELD_SERUM_SODIUM * log_MELD_BILIRUBIN + 
                    9.09 * log_MELD_INR + 
                    11.14 * log_MELD_SERUM_CREAT + 1.85 * MELD_ALBUMIN - 
                    1.83 * MELD_ALBUMIN * log_MELD_SERUM_CREAT + 6)
  ) %>% pull(MELD3)

preds_MELD_test1 = my_test1_df %>% 
  mutate(
    MELD3 = round(1.33 * ifelse(GENDER == "F", 1, 0) + 
                    4.56 * log_MELD_BILIRUBIN + 0.82 * MELD_SERUM_SODIUM - 
                    0.24 * MELD_SERUM_SODIUM * log_MELD_BILIRUBIN + 
                    9.09 * log_MELD_INR + 
                    11.14 * log_MELD_SERUM_CREAT + 1.85 * MELD_ALBUMIN - 
                    1.83 * MELD_ALBUMIN * log_MELD_SERUM_CREAT + 6)
  ) %>% pull(MELD3)

preds_MELD_test2 = my_test2_df %>% 
  mutate(
    MELD3 = round(1.33 * ifelse(GENDER == "F", 1, 0) + 
                    4.56 * log_MELD_BILIRUBIN + 0.82 * MELD_SERUM_SODIUM - 
                    0.24 * MELD_SERUM_SODIUM * log_MELD_BILIRUBIN + 
                    9.09 * log_MELD_INR + 
                    11.14 * log_MELD_SERUM_CREAT + 1.85 * MELD_ALBUMIN - 
                    1.83 * MELD_ALBUMIN * log_MELD_SERUM_CREAT + 6)
  ) %>% pull(MELD3)

preds_MELD_test3 = my_test3_df %>% 
  mutate(
    MELD3 = round(1.33 * ifelse(GENDER == "F", 1, 0) + 
                    4.56 * log_MELD_BILIRUBIN + 0.82 * MELD_SERUM_SODIUM - 
                    0.24 * MELD_SERUM_SODIUM * log_MELD_BILIRUBIN + 
                    9.09 * log_MELD_INR + 
                    11.14 * log_MELD_SERUM_CREAT + 1.85 * MELD_ALBUMIN - 
                    1.83 * MELD_ALBUMIN * log_MELD_SERUM_CREAT + 6)
  ) %>% pull(MELD3)
```

Re-fit MELD to obtain probabilities. 

```{r}
# Re-fit MELD model
MELD_fit = coxph(
  Surv(SurvTime_by90d, Death_by90d) ~ 
    GENDER + log_MELD_BILIRUBIN * MELD_SERUM_SODIUM + 
    log_MELD_INR + log_MELD_SERUM_CREAT * MELD_ALBUMIN, 
  data = train_df %>% 
    mutate(
      # Set lower/upper bounds for MELD variables and log transform
      MELD_BILIRUBIN = ifelse(INIT_BILIRUBIN < 1, 1, INIT_BILIRUBIN), 
      log_MELD_BILIRUBIN = log(MELD_BILIRUBIN), 
      MELD_SERUM_SODIUM = 137 - ifelse(INIT_SERUM_SODIUM > 137, 137, 
                                       ifelse(INIT_SERUM_SODIUM < 125, 125, INIT_SERUM_SODIUM)), 
      MELD_INR = ifelse(INIT_INR < 1, 1, INIT_INR), 
      log_MELD_INR = log(MELD_INR), 
      MELD_SERUM_CREAT = ifelse(INIT_SERUM_CREAT > 3 | INIT_DIALYSIS_PRIOR_WEEK == "Y", 3, 
                                ifelse(INIT_SERUM_CREAT < 1, 1, INIT_SERUM_CREAT)), 
      log_MELD_SERUM_CREAT = log(MELD_SERUM_CREAT), 
      MELD_ALBUMIN = 3.5 - ifelse(INIT_ALBUMIN > 3.5, 3.5, 
                                  ifelse(INIT_ALBUMIN < 1.5, 1.5, INIT_ALBUMIN))), 
  x = TRUE)

# Predict on data sets to get survival probabilities
preds_MELD_probs_train = as.vector(survfit(
  MELD_fit, 
  newdata = my_train_df)$surv[91,])
preds_MELD_probs_test1 = as.vector(survfit(
  MELD_fit, 
  newdata = my_test1_df)$surv[91,])
preds_MELD_probs_test2 = as.vector(survfit(
  MELD_fit, 
  newdata = my_test2_df)$surv[91,])
preds_MELD_probs_test3 = as.vector(survfit(
  MELD_fit, 
  newdata = my_test3_df)$surv[91,])
```

### Machine learning

#### Regularized Cox PH models (lasso, ridge, and elastic-net)

- Include all two-way interactions
- Tuning parameters:  regularization parameter `lambda`, plus weight for elastic-net `alpha` (`alpha` = 1 for lasso and `alpha` = 0 for ridge)
- Tuning method: built-in CV grid search (uses Harrel's C instead of default partial likelihood as CV loss)
- Currently using 1 SE instead of minimum lambda

```{r}
# Set up model matrices with all 2-way interactions
# Results in 241 predictors

my_train_mat = data.frame(
  my_train_df %>% 
    dplyr::select(SurvTime_by90d, Death_by90d), 
  model.matrix(Surv(SurvTime_by90d, Death_by90d) ~ .^2, # All 2-way interactions
               data = my_train_df %>% dplyr::select(-unlist(my_factor_vars)))
) %>% 
  # Exclude the garbage interactions
  dplyr::select(-c(X.Intercept., 
                   log_MELD_BILIRUBIN.INIT_BILIRUBIN, 
                   MELD_SERUM_SODIUM.INIT_SERUM_SODIUM, 
                   log_MELD_INR.INIT_INR, 
                   log_MELD_SERUM_CREAT.INIT_SERUM_CREAT, 
                   MELD_ALBUMIN.INIT_ALBUMIN, 
                   log_AGE.INIT_AGE))

my_test1_mat = data.frame(
  my_test1_df %>% 
    dplyr::select(SurvTime_by90d, Death_by90d), 
  model.matrix(Surv(SurvTime_by90d, Death_by90d) ~ .^2, # All 2-way interactions
               data = my_test1_df %>% dplyr::select(-unlist(my_factor_vars)))
) %>% 
  # Exclude the garbage interactions
  dplyr::select(-c(X.Intercept., 
                   log_MELD_BILIRUBIN.INIT_BILIRUBIN, 
                   MELD_SERUM_SODIUM.INIT_SERUM_SODIUM, 
                   log_MELD_INR.INIT_INR, 
                   log_MELD_SERUM_CREAT.INIT_SERUM_CREAT, 
                   MELD_ALBUMIN.INIT_ALBUMIN, 
                   log_AGE.INIT_AGE))

my_test2_mat = data.frame(
  my_test2_df %>% 
    dplyr::select(SurvTime_by90d, Death_by90d), 
  model.matrix(Surv(SurvTime_by90d, Death_by90d) ~ .^2, # All 2-way interactions
               data = my_test2_df %>% dplyr::select(-unlist(my_factor_vars)))
) %>% 
  # Exclude the garbage interactions
  dplyr::select(-c(X.Intercept., 
                   log_MELD_BILIRUBIN.INIT_BILIRUBIN, 
                   MELD_SERUM_SODIUM.INIT_SERUM_SODIUM, 
                   log_MELD_INR.INIT_INR, 
                   log_MELD_SERUM_CREAT.INIT_SERUM_CREAT, 
                   MELD_ALBUMIN.INIT_ALBUMIN, 
                   log_AGE.INIT_AGE))

my_test3_mat = data.frame(
  my_test3_df %>% 
    dplyr::select(SurvTime_by90d, Death_by90d), 
  model.matrix(Surv(SurvTime_by90d, Death_by90d) ~ .^2, # All 2-way interactions
               data = my_test3_df %>% dplyr::select(-unlist(my_factor_vars)))
) %>% 
  # Exclude the garbage interactions
  dplyr::select(-c(X.Intercept., 
                   log_MELD_BILIRUBIN.INIT_BILIRUBIN, 
                   MELD_SERUM_SODIUM.INIT_SERUM_SODIUM, 
                   log_MELD_INR.INIT_INR, 
                   log_MELD_SERUM_CREAT.INIT_SERUM_CREAT, 
                   MELD_ALBUMIN.INIT_ALBUMIN, 
                   log_AGE.INIT_AGE))
```

```{r, eval = FALSE}
# CV to find optimal lambda and alpha parameters for regularized Cox PH models
# ~30 minutes
set.seed(1)
fit_cv_enet = 
  cva.glmnet(Surv(SurvTime_by90d, Death_by90d) ~ ., 
             data = my_train_mat %>% filter(SurvTime_by90d > 0), 
             family = "cox", type.measure = "C")
```

```{r, echo = FALSE}
load("glmnet.rData")
```

```{r}
# Organize CV results
cv_res_tab = do.call(
  rbind, 
  lapply(1:length(fit_cv_enet$alpha), function(i) {
    data.frame(Alpha = fit_cv_enet$alpha[i], 
               as.data.frame(print(fit_cv_enet$modlist[[i]]))["1se",])
    })
)
rownames(cv_res_tab) = NULL
cv_res_tab = cv_res_tab[c(1, which.max(cv_res_tab$Measure), nrow(cv_res_tab)),] 
cv_res_tab = data.frame(Model = c("Ridge", "Elastic-net", "Lasso"), cv_res_tab)
cv_res_tab
```

```{r, eval = FALSE, message = FALSE}
# Predict on data sets
# Ridge
preds_ridge_train = as.numeric(
  predict(fit_cv_enet, my_train_mat, type = "response", 
          alpha = 0, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==0])
)
preds_ridge_test1 = as.numeric(
  predict(fit_cv_enet, my_test1_mat, type = "response", 
          alpha = 0, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==0])
)
preds_ridge_test2 = as.numeric(
  predict(fit_cv_enet, my_test2_mat, type = "response", 
          alpha = 0, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==0])
)
preds_ridge_test3 = as.numeric(
  predict(fit_cv_enet, my_test3_mat, type = "response", 
          alpha = 0, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==0])
)

# Elastic-net
enet_alpha = cv_res_tab$Alpha[cv_res_tab$Model == "Elastic-net"]
preds_enet_train = as.numeric(
  predict(fit_cv_enet, my_train_mat, type = "response", 
          alpha = enet_alpha,
          lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==enet_alpha])
)
preds_enet_test1 = as.numeric(
  predict(fit_cv_enet, my_test1_mat, type = "response", 
          alpha = enet_alpha, 
          lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==enet_alpha])
)
preds_enet_test2 = as.numeric(
  predict(fit_cv_enet, my_test2_mat, type = "response", 
          alpha = enet_alpha, 
          lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==enet_alpha])
)
preds_enet_test3 = as.numeric(
  predict(fit_cv_enet, my_test3_mat, type = "response", 
          alpha = enet_alpha, 
          lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==enet_alpha])
)

# Lasso
preds_lasso_train = as.numeric(
  predict(fit_cv_enet, my_train_mat, type = "response", 
          alpha = 1, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==1])
)
preds_lasso_test1 = as.numeric(
  predict(fit_cv_enet, my_test1_mat, type = "response", 
          alpha = 1, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==1])
)
preds_lasso_test2 = as.numeric(
  predict(fit_cv_enet, my_test2_mat, type = "response", 
          alpha = 1, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==1])
)
preds_lasso_test3 = as.numeric(
  predict(fit_cv_enet, my_test3_mat, type = "response", 
          alpha = 1, lambda = cv_res_tab$Lambda[cv_res_tab$Alpha==1])
)

save(fit_cv_enet, 
     preds_ridge_train, preds_ridge_test1, preds_ridge_test2, preds_ridge_test3, 
     preds_enet_train, preds_enet_test1, preds_enet_test2, preds_enet_test3, 
     preds_lasso_train, preds_lasso_test1, preds_lasso_test2, preds_lasso_test3, 
     file = "glmnet.rData")
```

#### Survival random forests

- Tuning parameters: minimum size of terminal node `nodesize` and number of variables to possibly split at each node `mtry`(other parameters include number of trees `ntree` and number of random splits `nsplit`)
- Tuning method: built-in grid search using out-of-sample error
- Solution expected to be approximately optimal. Increasing `sampsize` and/or `ntreeTry` should help? 

```{r, eval = FALSE}
# Setting doBest = TRUE returns a fitted random forest, but can't seem to use 
# it to get predictions
# 1 hour + 20 minutes to run (including refitting and predicting)
set.seed(1)
tune_rfsrc = randomForestSRC::tune(
  Surv(SurvTime_by90d, Death_by90d) ~ ., 
  data = my_train_df %>%
    dplyr::select(-names(my_factor_vars)), 
  sampsize = 0.3 * nrow(my_train_df), ntreeTry = 200)

# Fit RandomForestSRC survival forest
fit_rfsrc = rfsrc(Surv(SurvTime_by90d, Death_by90d) ~ ., 
                  data = my_train_df %>%
                    dplyr::select(-names(my_factor_vars)), 
                  nodesize = tune_rfsrc$optimal["nodesize"], 
                  mtry = tune_rfsrc$optimal["mtry"])

# Predict on data sets
preds_rfsrc_train = predict(fit_rfsrc)
preds_rfsrc_test1 = predict(fit_rfsrc, 
                            newdata = my_test1_df %>%
                              dplyr::select(-names(my_factor_vars)))
preds_rfsrc_test2 = predict(fit_rfsrc, 
                            newdata = my_test2_df %>%
                              dplyr::select(-names(my_factor_vars)))
preds_rfsrc_test3 = predict(fit_rfsrc, 
                            newdata = my_test3_df %>%
                              dplyr::select(-names(my_factor_vars)))

# Save model and predictions
save(tune_rfsrc, 
     fit_rfsrc, 
     preds_rfsrc_train, 
     preds_rfsrc_test1, 
     preds_rfsrc_test2, 
     preds_rfsrc_test3, 
     file = "rfsrc.rData")
```

```{r, echo = FALSE}
load("rfsrc.rData")
```

```{r}
# Default parameters: nodesize 15, mtry = sqrt(9-2) = 2-3, ntree = 500, nsplit = 10
tune_rfsrc$optimal
```

#### Gradient boosting models

GBM

- Tuning parameters: maximum depth of each tree `interaction.depth` and total number of trees `n.trees` (other parameters include minimum number of observations in the terminal nodes `n.minobsinnode` and learning rate `shrinkage`)
- Tuning method: CV grid search (grid adapted from `caret` package default to include default paramterization)

GLMBoost (currently NOT RUN: takes too long)

- Tuning parameters: number of initial boosting iterations `mstop` (other parameters include step size/shrinkage `nu`)
- Tuning method: CV grid search (grid adapted from `caret` package default to include default paramterization)

XGBoost-Linear

- Tuning parameters (out of many): L2 regularization `lambda`, L1 regularization `alpha`, and total number of trees `nrounds`
- Tuning method: CV grid search (grid adapted from `caret` package default to include default paramterization)
    
XGBoost-Tree

- Tuning parameters (out of many): maximum tree depth `max_depth`, total number of trees `nrounds`, step size shrinkage `eta`, subsample ratio of columns when constructing each tree `colsample_bytree`, and subsample ratio of the training instances `subsample`
- Tuning method: CV grid search (grid adapted from `caret` package default to include default paramterization)

```{r, eval = FALSE}
# Fit auto-tuners to training data
# GBM
# ~4 minutes
# getModelInfo("gbm")[[1]]$grid # caret default grid
gbm_at = auto_tuner(
  learner = lrn("surv.gbm"), 
  search_space = ps(
    interaction.depth = p_int(lower = 1, upper = 3), # Default 1
    n.trees = p_int(lower = 1*50, upper = 3*50) # Default 100
  ),
  resampling = rsmp("cv", folds = 10),
  measure = msr("surv.cindex"), 
  tuner = tnr("grid_search", resolution = 5))
set_seed(100)
gbm_at$train(task_train)

# GLMBoost
# getModelInfo("glmboost")[[1]]$grid # caret default grid
# glmboost_at = auto_tuner(
#   learner = lrn("surv.glmboost"), 
#   search_space = ps(
#     mstop =  p_int(lower = 1*50, upper = 3*50) # Default 100
#   ),
#   resampling = rsmp("cv", folds = 10),
#   measure = msr("surv.cindex"), 
#   tuner = tnr("grid_search", resolution = 5))
# set_seed(100)
# glmboost_at$train(task_train)

# XGBoost-Linear
# ~5 minutes
# getModelInfo("xgbLinear")[[1]]$grid # caret default grid
xgblinear_at = auto_tuner(
  learner = lrn("surv.xgboost", booster = "gblinear"), 
  search_space = ps(
    lambda = p_dbl(lower = 0, upper = 10^-1), # Default 0
    alpha = p_dbl(lower = 0, upper = 10^-1), # Default 0
    nrounds = p_int(lower = 1*50, upper = 3*50) # No default
  ),
  resampling = rsmp("cv", folds = 10),
  measure = msr("surv.cindex"), 
  tuner = tnr("grid_search", resolution = 5))
set_seed(100)
xgblinear_at$train(task_train)

# XGBoost-Tree
# getModelInfo("xgbTree")[[1]]$grid # caret default grid
# This takes a long time because there are so many hyperparmaeters, so 
# consider limiting parameters or using a lower resolution? 
# 9.5 hours
# 1:17am
xgbtree_at = auto_tuner(
  learner = lrn("surv.xgboost", booster = "gbtree"), 
  search_space = ps(
    max_depth = p_int(lower = 1, upper = 6), # Default 6
    nrounds = p_int(lower = 1*50, upper = 3*50), # No default
    eta = p_dbl(lower = 0.3, upper = 0.4), # Default 0.3
    colsample_bytree =  p_dbl(lower = 0.6, upper = 1), # Default 1
    subsample = p_dbl(lower = 0.5, upper = 1) # Default 1
  ),
  resampling = rsmp("cv", folds = 10),
  measure = msr("surv.cindex"), 
  tuner = tnr("grid_search", resolution = 5))
set_seed(100)
xgbtree_at$train(task_train)

# List of auto-tuned models
mods_boost_at = list(
  "gbm" = gbm_at, 
  #"glmboost" = glmboost_at, 
  "xgblinear" = xgblinear_at, 
  "xgbtree" = xgbtree_at
)

# Re-fit tuned models (necessary to do it this way to get survival probabilities)
set.seed(1)
fit_boost = lapply(mods_boost_at, function(at) {
  my_learn = ppl("crankcompositor", 
                 ppl("distrcompositor", at$learner),
                 response = TRUE, overwrite = FALSE, 
                 method = "mean", graph_learner = TRUE)
  my_learn$train(task_train)
})
names(fit_boost) = names(mods_boost_at)

# Predict on data sets (linear predictor seems to discriminate better than 
# 90-day survival probability)
preds_boost_train = sapply(names(mods_boost_at), function(fun) {
  # fit_boost[[fun]]$predict(task_train)$distr$survival(90)
  fit_boost[[fun]]$predict(task_train)$lp
})
preds_boost_test1 = sapply(names(mods_boost_at), function(fun) {
  # fit_boost[[fun]]$predict(task_test1)$distr$survival(90)
  fit_boost[[fun]]$predict(task_test1)$lp
})
preds_boost_test2 = sapply(names(mods_boost_at), function(fun) {
  # fit_boost[[fun]]$predict(task_test2)$distr$survival(90)
  fit_boost[[fun]]$predict(task_test2)$lp
})
preds_boost_test3 = sapply(names(mods_boost_at), function(fun) {
  # fit_boost[[fun]]$predict(task_test3)$distr$survival(90)
  fit_boost[[fun]]$predict(task_test3)$lp
})

save(gbm_at, #glmboost_at, 
     xgblinear_at, xgbtree_at, 
     preds_boost_train, preds_boost_test1, 
     preds_boost_test2, preds_boost_test3, 
     file = "boost_at.rData")
```

```{r, echo = FALSE}
load("boost_at.rData")
```

```{r}
gbm_at$tuning_result %>% 
  dplyr::select(interaction.depth:n.trees, surv.cindex)
# glmboost_at$tuning_result  %>% 
#   dplyr::select(mstop, surv.cindex)
xgblinear_at$tuning_result %>% 
  dplyr::select(lambda:nrounds, surv.cindex)
xgbtree_at$tuning_result %>% 
  dplyr::select(max_depth:subsample, surv.cindex)
```

CoxBoost

- Tuning parameters: penalty at each step `penalty` and number of boosting steps`stepno`
- Tuning method: coarse line search to find the optimal penalty and CV to find the optimal number of boosting steps

```{r, eval = FALSE}
# Run CV to identify optimal penalty and number of boosting steps
# Takes ~10.5 hours and exits with warning that 
# "Exceeded iter.max in search for penalty parameter" 
set.seed(1)
fit_cv_coxboost = optimCoxBoostPenalty(
  status = my_train_df$Death_by90d, 
  time = my_train_df$SurvTime_by90d, 
  x = model.matrix(~., my_train_df %>%
                     dplyr::select(-unlist(my_factor_vars)))[,-1], 
  parallel = TRUE)

# Re-fit optimal model
fit_coxboost = CoxBoost(
  status = my_train_df$Death_by90d, 
  time = my_train_df$SurvTime_by90d, 
  x = model.matrix(~., my_train_df %>%
                     dplyr::select(-unlist(my_factor_vars)))[,-1], 
  stepno = fit_cv_coxboost$cv.res$optimal.step,
  penalty = fit_cv_coxboost$penalty) 

# Predict 90-day survival for data sets
# For some reason, risk/CIF gets you all 1s and 0s
preds_coxboost_train = as.numeric(predict(
  fit_coxboost, 
  newstatus = my_train_df$Death_by90d, 
  newtime = my_train_df$SurvTime_by90d, 
  newdata = model.matrix(~., my_train_df %>%
                           dplyr::select(-unlist(my_factor_vars)))[,-1], 
  type = "CIF", times = 90))
preds_coxboost_test1 = as.numeric(predict(
  fit_coxboost, 
  newstatus = my_test1_df$Death_by90d, 
  newtime = my_test1_df$SurvTime_by90d, 
  newdata = model.matrix(~., my_test1_df %>%
                           dplyr::select(-unlist(my_factor_vars)))[,-1], 
  type = "CIF", times = 90))
preds_coxboost_test2 = as.numeric(predict(
  fit_coxboost, 
  newstatus = my_test2_df$Death_by90d, 
  newtime = my_test2_df$SurvTime_by90d, 
  newdata = model.matrix(~., my_test2_df %>%
                           dplyr::select(-unlist(my_factor_vars)))[,-1], 
  type = "CIF", times = 90))
preds_coxboost_test3 = as.numeric(predict(
  fit_coxboost, 
  newstatus = my_test3_df$Death_by90d, 
  newtime = my_test3_df$SurvTime_by90d, 
  newdata = model.matrix(~., my_test3_df %>%
                           dplyr::select(-unlist(my_factor_vars)))[,-1], 
  type = "CIF", times = 90))

save(fit_cv_coxboost, fit_coxboost, 
     preds_coxboost_train, preds_coxboost_test1, 
     preds_coxboost_test2, preds_coxboost_test3, 
     file = "coxboost.rData")
```

```{r, echo = FALSE}
load("coxboost.rData")
```

```{r}
c(penalty = fit_cv_coxboost$penalty, 
  "number of steps" = fit_cv_coxboost$cv.res$optimal.step)
```

#### Neural networks (CoxTime, DeepHit, DeepSurv, logistic hazard, and PC hazard)

- Tuning parameters (out of many): dropout rate `dropout`, weight decay `weight_decay`, learning rate `learning_rate`, and number of nodes `num_nodes` in a layer/number of layers
- Tuning method: random search + additional CV to choose between "optimal" and default models

```{r, eval = FALSE}
# Set up search space
search_space = ps(
  dropout = p_dbl(lower = 0, upper = 1),
  weight_decay = p_dbl(lower = 0, upper = 0.5),
  learning_rate = p_dbl(lower = 0, upper = 1),
  nodes = p_int(lower = 1, upper = 32),
  k = p_int(lower = 1, upper = 4)
)
search_space$trafo <- function(x, param_set) {
  x$num_nodes = rep(x$nodes, x$k)
  x$nodes = x$k = NULL
  return(x)
}


# Fit auto-tuners to training data
# ~1.5 hours collectively (including benchmarking), with large variation 
# between autotuners for models (DeepSurv is short)
# CoxTime 
coxtime_at = auto_tuner(
  learner = lrn("surv.coxtime", frac = 0.3, early_stopping = TRUE, 
                epochs = 100, optimizer = "adam"), 
  search_space = search_space,
  resampling = rsmp("cv", folds = 3),
  measure = msr("surv.cindex"), 
  terminator = trm("evals", n_evals = 60),
  tuner = tnr("random_search"))
set_seed(100)
coxtime_at$train(task_train)

# DeepHit 
deephit_at = auto_tuner(
  learner = lrn("surv.deephit", frac = 0.3, early_stopping = TRUE, 
                epochs = 100, optimizer = "adam"), 
  search_space = search_space,
  resampling = rsmp("cv", folds = 3),
  measure = msr("surv.cindex"), 
  terminator = trm("evals", n_evals = 60),
  tuner = tnr("random_search"))
set_seed(108)
deephit_at$train(task_train)

# DeepSurv 
deepsurv_at = auto_tuner(
  learner = lrn("surv.deepsurv", frac = 0.3, early_stopping = TRUE, 
                epochs = 100, optimizer = "adam"), 
  search_space = search_space,
  resampling = rsmp("cv", folds = 3),
  measure = msr("surv.cindex"), 
  terminator = trm("evals", n_evals = 60),
  tuner = tnr("random_search"))
set_seed(100)
deepsurv_at$train(task_train)

# Logistic hazard 
loghaz_at = auto_tuner(
  learner = lrn("surv.loghaz", frac = 0.3, early_stopping = TRUE, 
                epochs = 100, optimizer = "adam"), 
  search_space = search_space,
  resampling = rsmp("cv", folds = 3),
  measure = msr("surv.cindex"), 
  terminator = trm("evals", n_evals = 60),
  tuner = tnr("random_search"))
set_seed(100) 
loghaz_at$train(task_train)

# PC hazard 
pchazard_at = auto_tuner(
  learner = lrn("surv.pchazard", frac = 0.3, early_stopping = TRUE, 
                epochs = 100, optimizer = "adam"), 
  search_space = search_space,
  resampling = rsmp("cv", folds = 3),
  measure = msr("surv.cindex"), 
  terminator = trm("evals", n_evals = 60),
  tuner = tnr("random_search"))
set_seed(100)
pchazard_at$train(task_train)

# Run CV benchmark to compare tuned and default models
learners = list(
  lrn("surv.coxtime", id = "surv.coxtime.default"), 
  coxtime_at$learner, 
  lrn("surv.deephit", id = "surv.deephit.default"),
  deephit_at$learner, 
  lrn("surv.deepsurv", id = "surv.deepsurv.default"),
  deepsurv_at$learner, 
  lrn("surv.loghaz", id = "surv.loghaz.default"),
  loghaz_at$learner, 
  lrn("surv.pchazard", id = "surv.pchazard.default"),
  pchazard_at$learner
)
set.seed(1)
design = benchmark_grid(task_train, learners, rsmp("cv", folds = 5))
bm = benchmark(design)
```

```{r, echo = FALSE}
load("nn_at.rData")
```

```{r}
# List of auto-tuned models
mods_at = list(
  "coxtime" = coxtime_at, 
  "deephit" = deephit_at, 
  "deepsurv" = deepsurv_at, 
  "loghaz" = loghaz_at, 
  "pchazard" = pchazard_at
)

# Default parameters: dropout = 0.1, weight_decay = 0, learning_rate = 0.01, 
# num_nodes = c(32, 32)
# Tuned parameters
sapply(mods_at, function(x) {
  x$tuning_result %>% 
  dplyr::select(dropout:k, surv.cindex)
}) %>% t()

# Summary of C-statistics for default and best tuned model
bm_df = data.frame(
  Model = rep(c("coxtime", "deephit", "deepsurv", "loghaz", "pchazard"), each = 2), 
  Parameters = rep(c("default", "tuned"), times = 5), 
  surv.cindex = bm$aggregate(msr("surv.cindex"))$surv.cindex
) %>% 
  rename("CV C-index" = surv.cindex) 
```

```{r, eval = FALSE}
# pycox models to try
pycox_mods = c(
  # Based on the Cox PH with time-varying effects
  "coxtime" = coxtime, 
  # Based on the PMF of a discrete Cox model
  "deephit" = deephit, 
  # Based on the partial likelihood from a Cox PH
  "deepsurv" = deepsurv, 
  # Discrete neural networks based on a cross-entropy loss and predictions of a 
  # discrete hazard function, also known as Nnet-Survival
  # Logistic hazard
  "loghaz" = loghaz, 
  # PC hazard
  "pchazard" = pchazard)

# Re-fit the best of each pycox model (either default or tuned)
set_seed(1)
fit_pycox = lapply(names(pycox_mods), function(fun) {
  param_spec = bm_df$Parameters[which.max(bm_df$`CV C-index`[bm_df$Model == fun])]
  if (param_spec == "default") {
    pycox_mods[[fun]](Surv(SurvTime_by90d, Death_by90d) ~ ., 
                      data = my_train_df %>%
                        dplyr::select(-names(my_factor_vars)))
  } else {
    tune_res = mods_at[[fun]]$tuning_result
    pycox_mods[[fun]](Surv(SurvTime_by90d, Death_by90d) ~ ., 
                      data = my_train_df %>%
                        dplyr::select(-names(my_factor_vars)), 
                      dropout = tune_res$dropout, 
                      num_nodes = rep(tune_res$nodes, tune_res$k), 
                      weight_decay = tune_res$weight_decay, 
                      learning_rate = tune_res$learning_rate, 
                      frac = 0.3, early_stopping = TRUE, 
                      epochs = 100, optimizer = "adam")
  }
})
names(fit_pycox) = names(pycox_mods)

# Predict on data sets (90-day survival probability)
preds_pycox_train = sapply(names(pycox_mods), function(fun) {
  predict(fit_pycox[[fun]], type = "survival")[,"90"]
})
preds_pycox_test1 = sapply(names(pycox_mods), function(fun) {
  predict(fit_pycox[[fun]], 
          newdata = my_test1_df %>%
            dplyr::select(-names(my_factor_vars)), 
          type = "survival")[,"90"]
})
preds_pycox_test2 = sapply(names(pycox_mods), function(fun) {
  predict(fit_pycox[[fun]], 
          newdata = my_test2_df %>%
            dplyr::select(-names(my_factor_vars)), 
          type = "survival")[,"90"]
})
preds_pycox_test3 = sapply(names(pycox_mods), function(fun) {
  predict(fit_pycox[[fun]], 
          newdata = my_test3_df %>%
            dplyr::select(-names(my_factor_vars)), 
          type = "survival")[,"90"]
})

# Save model and predictions
save(coxtime_at, deephit_at, deepsurv_at, loghaz_at, pchazard_at, 
     bm, bm_df, fit_pycox, 
     preds_pycox_train, preds_pycox_test1, 
     preds_pycox_test2, preds_pycox_test3, 
     file = "nn_at.rData")
```

## Concordance

Calculate the concordance for each model and dataset. 

```{r}
# Put all predictions together
all_preds_train = data.frame(
  meld3 = preds_MELD_train, 
  ridge = preds_ridge_train, 
  enet = preds_enet_train, 
  lasso = preds_lasso_train, 
  rfsrc = preds_rfsrc_train$predicted.oob, 
  preds_boost_train, 
  coxboost = preds_coxboost_train,
  1 - preds_pycox_train)
all_preds_test1 = data.frame(
  meld3 = preds_MELD_test1, 
  ridge = preds_ridge_test1, 
  enet = preds_enet_test1, 
  lasso = preds_lasso_test1, 
  rfsrc = preds_rfsrc_test1$predicted, 
  preds_boost_test1, 
  coxboost = preds_coxboost_test1,
  1 - preds_pycox_test1)
all_preds_test2 = data.frame(
  meld3 = preds_MELD_test2, 
  ridge = preds_ridge_test2, 
  enet = preds_enet_test2, 
  lasso = preds_lasso_test2, 
  rfsrc = preds_rfsrc_test2$predicted, 
  preds_boost_test2, 
  coxboost = preds_coxboost_test2,
  1 - preds_pycox_test2)
all_preds_test3 = data.frame(
  meld3 = preds_MELD_test3, 
  ridge = preds_ridge_test3, 
  enet = preds_enet_test3, 
  lasso = preds_lasso_test3, 
  rfsrc = preds_rfsrc_test3$predicted, 
  preds_boost_test3, 
  coxboost = preds_coxboost_test3,
  1 - preds_pycox_test3)

# Concordance
all_concordances = as.data.frame(cbind(
  "Training" = sapply(names(all_preds_train), function(x) {
    concordance(Surv(SurvTime_by90d, Death_by90d) ~ all_preds_train[,x], 
                data = my_train_df, reverse = TRUE)$concordance
  }), 
  "Test 1" = sapply(names(all_preds_test1), function(x) {
    concordance(Surv(SurvTime_by90d, Death_by90d) ~ all_preds_test1[,x], 
                data = my_test1_df, reverse = TRUE)$concordance
  }), 
  "Test 2" = sapply(names(all_preds_test2), function(x) {
    concordance(Surv(SurvTime_by90d, Death_by90d) ~ all_preds_test2[,x], 
                data = my_test2_df, reverse = TRUE)$concordance
  }), 
  "Test 3" = sapply(names(all_preds_test3), function(x) {
    concordance(Surv(SurvTime_by90d, Death_by90d) ~ all_preds_test3[,x], 
                data = my_test3_df, reverse = TRUE)$concordance
  })
))
round(all_concordances, 5)
```

Test if the concordances are significantly different for MELD 3.0 and each machine learning model. 

```{r, eval = FALSE}
# Compare C-indices
# 45 minutes
all_concordance_pvals = as.data.frame(cbind(
  "Training" = sapply(all_preds_train[,-1], function(x) {
    compareC(my_train_df$SurvTime_by90d, my_train_df$Death_by90d, 
             preds_MELD_train, x)$pval
  }), 
  "Test 1" = sapply(all_preds_test1[,-1], function(x) {
    compareC(my_test1_df$SurvTime_by90d, my_test1_df$Death_by90d, 
             preds_MELD_test1, x)$pval
  }), 
  "Test 2" = sapply(all_preds_test2[,-1], function(x) {
    compareC(my_test2_df$SurvTime_by90d, my_test2_df$Death_by90d, 
             preds_MELD_test2, x)$pval
  }), 
  "Test 3" = sapply(all_preds_test3[,-1], function(x) {
    compareC(my_test3_df$SurvTime_by90d, my_test3_df$Death_by90d, 
             preds_MELD_test3, x)$pval
  })
))
save(all_concordances, all_concordance_pvals, file = "all_concordance_pvals.rData")
```

```{r, echo = FALSE}
load("all_concordance_pvals.rData")
```

```{r}
round(all_concordance_pvals, 5)
```

## Tables

### Table 1

Median (IQR) reported for continuous variables and count (percent) for binary/categorical variables. 

```{r}
# Stack all data and pull out variables of interest
tab1_df = rbind(
  train_df %>% dplyr::select(-DGN_OSTXT_TCR,-DGN2_OSTXT_TCR) %>% mutate(Data = "Training"), 
  test1_df %>% dplyr::select(-DGN_OSTXT_TCR,-DGN2_OSTXT_TCR) %>% mutate(Data = "Test 1"), 
  test2_df %>% mutate(Data = ifelse(mdy(INIT_DATE) < as.Date("2021-01-01"), 
                                    "Test 2", "Test 3")) 
) %>% 
  mutate(
    Data = factor(Data, levels = c("Training", paste("Test", 1:3))), 
    Female = ifelse(GENDER == "F", 1, 0), 
    dialysis = ifelse(INIT_DIALYSIS_PRIOR_WEEK == "Y", 1, 0), 
    ethnicity_cat = fct_relabel(factor(Ethnicity_Cat), 
                                ~str_sub(., start = 3)), 
    enceph = fct_recode(factor(INIT_ENCEPH), 
                        None = "1", "1-2" = "2", "3-4" = "3"), 
    dgn = factor(case_when(
      DGN_TCR %in% as.character(c(4100:4107, 4201:4207, 4216)) ~ 
        "Viral hepatitis", 
      DGN_TCR %in% as.character(c(4215, 4217:4219)) ~ 
        "Alcohol associated liver disease", 
      DGN_TCR %in% "4214" ~ "NASH", 
      DGN_TCR %in% as.character(c(4220, 4230, 4231, 4235, 4240:4242, 4245, 
                                  4250, 4255, 4260, 4264, 4270:4272, 4275)) ~ 
        "Cholestatic", 
      TRUE ~ "Other"), 
      levels = c("Alcohol associated liver disease", "Cholestatic", "NASH", 
                 "Viral hepatitis", "Other"))) %>% 
  dplyr::select(Data, 
                "Age" = INIT_AGE, 
                Female, 
                "Race/ethnicity" = ethnicity_cat, 
                "Primary diagnosis" = dgn, 
                Bilirubin = INIT_BILIRUBIN, 
                "Serum sodium" = INIT_SERUM_SODIUM, 
                INR = INIT_INR, 
                "Serum creatinine" = INIT_SERUM_CREAT, 
                "Albumin" = INIT_ALBUMIN, 
                "Dialysis" = dialysis,
                "Ascites" = ASCITES_BL, 
                "Encephalopathy" = enceph, 
                "Dead at 90 days" = Death_by90d)

# Create Table 1
tab1_gt = tab1_df %>% 
  tbl_summary(by = Data) %>%
  modify_footnote(everything() ~ NA)

as_gt(tab1_gt) %>% gt::gtsave(filename = "tab1.docx")
tab1_gt
```

### Table 2

We define the models fit using only MELD variables as one family and the models that incorporate additional variables as a second family. Each family has a total of (13 models) x (2 test sets) = 26 tests, so the Bonferroni threshold is 0.05/26 = 0.001923. 

```{r}
# Load C-statistics and p-values from models that use only MELD variables
all_concordances2 = all_concordances
all_concordance_pvals2 = all_concordance_pvals
load("../2024.06.26/all_concordance_pvals.rData")

# Put all C-statistics and p-values together
tab2_df = data.frame(
  Variables = c("MELD 3.0 components only", 
                rep("", nrow(all_concordances) - 1), 
                "MELD 3.0 components + additional", 
                rep("", nrow(all_concordances) - 2)), 
  Model = c("MELD 3.0", 
            rep(c("Ridge", "Elastic-net", "Lasso", 
                  "Random forest", "GBM", 
                  "XGBoost-Linear", "XGBoost-Tree", "CoxBoost", 
                  "Cox-Time", "DeepHit", "DeepSurv", 
                  "Logistic-Hazard", "PC-Hazard"), times = 2)), 
  rbind(cbind(all_concordances, 
              rbind(meld3 = NA, all_concordance_pvals)), 
        cbind(all_concordances2[-1,], 
              all_concordance_pvals2))
)[, c(1:2, 2 + rep(1:ncol(all_concordances), each = 2) + 
                    rep(c(0, ncol(all_concordances)), 
                        times = ncol(all_concordances)))] %>% 
  # Round to 3rd decimal place and replace those <0.001 with "<0.001"
  mutate_if(is.numeric, ~as.character(replace_na(
    ifelse(. < 0.001, "<0.001", round(., 3)), "Reference"))) %>% 
  `rownames<-`( NULL )
# Nice names
tab2_df = rbind(
  c("", "", rep(c("C-statistic", "p-value"), ncol(all_concordances))), 
  tab2_df
)
names(tab2_df) = c("Variables", "Model", 
                   "Training", "", "Test 1", "", 
                   "Test 2", "", "Test 3", "")

write.csv(tab2_df, file = "tab2.csv", row.names = FALSE)
kbl(tab2_df) %>% # Align columns right
  # This table style puts horizontal lines at the top and bottom
  kable_classic(html_font = "\"Source Sans Pro\", helvetica, sans-serif") %>% 
  # Bold headers
  row_spec(0:1, bold = TRUE) %>%
  # Extra horizontal lines
  row_spec(c(1, 15), extra_css = "border-bottom: 1px solid;")
```
