---
title: "Tables 1-2"
output: html_document
date: "2024-09-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(tidyverse)
library(kableExtra)
library(gtsummary)

# Set background to be white for all ggplots
theme_set(theme_classic())
```

Read in the training and test data. 

```{r}
dir = "/Users/jliang/Library/CloudStorage/Box-Box/MELD\ QSU/data/"
train_df = read.csv(paste0(dir, "MELD3_Dev_Cohort.csv"))
test_df = read.csv("/Users/jliang/Library/CloudStorage/Box-Box/STAR_files_as_of_4_5_2024/MELD\ 3.0/MELD3_SecVal_Cohort_2019_23.csv")
```

## Table 1

Median (IQR) reported for continuous variables and count (percent) for binary/categorical variables. 

```{r}
# Stack all data and pull out variables of interest
tab1_df = rbind(
  train_df %>% dplyr::select(-DGN_OSTXT_TCR,-DGN2_OSTXT_TCR) %>% mutate(Data = "Training"), 
  test_df %>% 
    # 1/1/2021 (inclusive) to end of available data
    filter(mdy(INIT_DATE) >= as.Date("2021-01-01")) %>% 
    mutate(Data = "Test") 
) %>% 
  mutate(
    Data = factor(Data, levels = c("Training", "Test")), 
    Female = ifelse(GENDER == "F", 1, 0), 
    dialysis = ifelse(INIT_DIALYSIS_PRIOR_WEEK == "Y", 1, 0), 
    ethnicity_cat = fct_relabel(factor(Ethnicity_Cat), 
                                ~str_sub(., start = 3)), 
    enceph = fct_recode(factor(INIT_ENCEPH), 
                        None = "1", "1-2" = "2", "3-4" = "3"), 
    dgn = factor(case_when(
      DGN_TCR %in% as.character(c(4100:4107, 4201:4207, 4216)) ~ 
        "Viral hepatitis", 
      DGN_TCR %in% as.character(c(4215, 4217:4219)) ~ 
        "Alcohol associated liver disease", 
      DGN_TCR %in% "4214" ~ "NASH", 
      DGN_TCR %in% as.character(c(4220, 4230, 4231, 4235, 4240:4242, 4245, 
                                  4250, 4255, 4260, 4264, 4270:4272, 4275)) ~ 
        "Cholestatic", 
      TRUE ~ "Other"), 
      levels = c("Alcohol associated liver disease", "Cholestatic", "NASH", 
                 "Viral hepatitis", "Other"))) %>% 
  dplyr::select(Data, 
                "Age" = INIT_AGE, 
                Female, 
                "Race/ethnicity" = ethnicity_cat, 
                "Primary diagnosis" = dgn, 
                Bilirubin = INIT_BILIRUBIN, 
                "Serum sodium" = INIT_SERUM_SODIUM, 
                INR = INIT_INR, 
                "Serum creatinine" = INIT_SERUM_CREAT, 
                "Albumin" = INIT_ALBUMIN, 
                "Dialysis" = dialysis,
                "Ascites" = ASCITES_BL, 
                "Encephalopathy" = enceph, 
                "Dead at 90 days" = Death_by90d)

# Create Table 1
tab1_gt = tab1_df %>% 
  tbl_summary(by = Data) %>%
  modify_footnote(everything() ~ NA)

as_gt(tab1_gt) %>% gt::gtsave(filename = "tab1.docx")
tab1_gt
```

## Table 2

We define the models/hypothesis tests run for each analysis and dataset as a family. Each family has a total of 13 tests, so the Bonferroni threshold is 0.05/13 = 0.00385.  

- Analysis 1: with biologically relevant transformations (closest to MELD 3.0)
- Analysis 2: no variable caps
- Analysis 3: no variable caps + additional clinical variables

```{r}
# Initialize lists for C-statistics and p-values
all_concordances_list = list()
all_concordance_pvals_list = list()
# Sub-directories to search for results
my_subdirs = c("analysis1", "analysis2", "analysis3")

# Load C-statistics and p-values from all analyses
for (i in 1:length(my_subdirs)) {
  load(paste0(my_subdirs[i], "/all_concordance_pvals.rData"))
  all_concordances_list[[i]] = all_concordances
  all_concordance_pvals_list[[i]] = all_concordance_pvals
}

# Put all C-statistics and p-values together
tab2_df = data.frame(
  Model = c("MELD 3.0", "Ridge", "Elastic-net", "Lasso", 
            "Random forest", "GBM", 
            "XGBoost-Linear", "XGBoost-Tree", "CoxBoost", 
            "Cox-Time", "DeepHit", "DeepSurv", "Logistic-Hazard", "PC-Hazard"), 
  cbind(cbind(all_concordances_list[[1]][,"Test"], 
              c(meld3 = NA, all_concordance_pvals_list[[1]][,"Test"])), 
        cbind(all_concordances_list[[2]][, "Test"], 
              c(meld3 = NA, all_concordance_pvals_list[[2]][,"Test"])), 
        cbind(all_concordances_list[[3]][,"Test"], 
              c(meld3 = NA, all_concordance_pvals_list[[3]][,"Test"])))
) %>% 
  # Round to 3rd decimal place and replace those <0.001 with "<0.001"
  mutate_if(is.numeric, ~as.character(replace_na(
    ifelse(. < 0.001, "<0.001", round(., 3)), "Reference"))) %>% 
  `rownames<-`( NULL )
# Nice names
tab2_df = rbind(
  c("", rep(c("C-statistic", "p-value"), length(my_subdirs))), 
  tab2_df
)
names(tab2_df) = c("Model", "Analysis 1", "", 
                   "Analysis 2", "", "Analysis 3", "")

write.csv(tab2_df, file = "tab2.csv", row.names = FALSE)
kbl(tab2_df) %>% # Align columns right
  # This table style puts horizontal lines at the top and bottom
  kable_classic(html_font = "\"Source Sans Pro\", helvetica, sans-serif") %>% 
  # Bold headers
  row_spec(0:1, bold = TRUE)
```
